Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.Rmd
Commit 2024-06-09 20:15

<!-- Aerosol change -->
The above clear sky reference values are based on the climatological AOD and WC,
hence they cannot describe accurately the long term variation of
$\text{GHI}_\text{ref}$ due to changes in the two atmospheric constituents, mainly
AOD.  As reported by @Natsis2023, there is a long term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data was
attributed to long-term changes in aerosol effects.

In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.

>>>>

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents, mainly AOD. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.
